---
title: "How landscape habitat influences wild eastern turkey populations within New York State"
author: Marissa Zimmer 
output:
  html_document:
      theme: cerulean
  
---

# Introduction

New York State’s (NYS) wild turkey population constitutes a significant part of conservation concerns and support for the hunting community within NYS. In the early 1950s, the Department of Conservation  (NYSDEC) helped restore and repopulate NYS wild turkey populations. The NYSDEC had also partnered with the National Wild Turkey Federation and SUNY Environmental Science and Forestry college in 2010 to research the impact of weather patterns and hunting zones on turkey populations. The results of this study have not been published at this time.

The aim of this project is to research, gather and compare data that involves understanding and explaining how land cover, tree canopy cover, elevation and precipitation can affect wild turkey population levels. Within NYS many individuals have noticed seeing less wild turkeys than in years past and changing habitat factors in localized areas may be one factor affecting population levels. By looking at these environmantal variables the aim of this project is to see if areas with less suitable environmental variables correlate to areas of lower population levels. If the results of this study come back to support this hypothesis then there is room for continued research into habitat suitability within NYS for wild turkey populations. 



# Materials and methods

For this project it was important to highlight the key environmental variables that affect wild turkey populations within New York State. Two of the environmental variables that were used, landcover and elevation were taken from the U.S. Geological Survey (USGS). Specifically, the data is a global digital elevation model (DEM) with a horizontal grid spacing of 30 arc seconds (approximately 1 kilometer) from several raster and vector sources of information. The USGS seperated the data into tiles on a world map which can be selected by grid. The data for elevation was completed in 1996 and was developed over a three year period. For this project I was able to select a grid tile that included NY. The next environmental variable that was chosen for this project was Tree canopy cover from the Global Land Analysis & Discovery. The global tree cover data are per pixel estimates of the 2010 percent maximum (peak growing season) tree canopy cover from cloud-free annual growing season composite Landset 7 ETM + data. A median from the annual tree canopy cover cover values from 2009-2011 were used to estimate the 2010 tree cover. Like the Landcover data the tree cover canopy data was chosen from a grid style map with a spatial resoultion of 1 arc-second per pixel, approximately 30 m per pixel. The last environmental variable, precipitation was taken from the PRISM Climate Group from Oregon State University. I selected the temporal period May of 2020. The month of May was chosen based on wild eastern Turkeys breeding times. During the month of May their breeding and nesting behaviors peak. 


# Step 1: Loading the required packages 


```{r, message=F, warning=F}

library(raster)
library(sf)
library(tigris)
library(rasterVis)
library(rgdal)
library(dismo)
library(ggplot2)
library(viridis)

# cache the results for quick compiling
options(tigris_use_cache = TRUE)
 
```


# Step 2: The Data

```{r, message=F, warning=F}
ebird_data <-read.csv(file = "C:/Users/maris/OneDrive/Desktop/maxent_turkey3.csv")

#Landcover from Dr. Williams 
land_cover = raster ("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/landcover_crop.asc")
crs(land_cover)=4326


#Treecover from Global Forest Watch
tree_cover = raster("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/Hansen_GFC-2019-v1.7_treecover2000_50N_080W.tif")


#Precipitation data from PRISM 
precip_cover = raster("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/PRISM_ppt_stable_4kmM3_2020_asc.asc")
crs(precip_cover)=4326


#Elevation data from Dr. Williams 
elevation_cover = raster ("C:/Users/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/elevation.tif")

```

let's take a quick look at all four environmental layers 

```{r, message=F, warning=F}
land_cover
tree_cover
precip_cover
elevation_cover
```

Now we need to make all the layers talk to each other 
I chose to have each layer contain the same properties as the landcover data 

```{r, message=F, warning=F}
ny_counties <-counties(state = "NY", cb = FALSE,
                       resolution = "500k", year = 2020) %>%
                        st_union() %>%
                         st_transform(crs=projection(land_cover))
#landcover 

land_coverny= crop(land_cover, as(ny_counties, "Spatial"))

#treecover

tree_coverny = crop(tree_cover, 
                    as(st_transform(ny_counties,st_crs(tree_cover)), "Spatial"))#%>%

tree_coverny2 = projectRaster(tree_coverny, to=land_coverny,method="ngb")

#elevation 
elevation_coverny= crop(elevation_cover,
                   as(st_transform(ny_counties,st_crs(elevation_cover)), "Spatial"))#%>%

elevation_coverny2 = projectRaster(elevation_coverny,to=land_coverny, method = "ngb")

#precipitation
precip_coverny = crop(precip_cover, 
                      as(st_transform(ny_counties,st_crs(precip_cover)), "Spatial"))#%>%

precip_coverny2 = projectRaster(precip_cover, to=land_coverny,method="ngb")

#extent time

extent_1 = extent(land_coverny)
precip_cover = crop(precip_cover, extent_1)
tree_coverny2 = crop(tree_coverny2, extent_1)
elevation_coverny2 = crop(elevation_coverny2, extent_1)
```

Let's relook at all of the environmental layers again and make sure they match:

```{r, message=F, warning=F}
land_coverny
tree_coverny2
precip_coverny2
elevation_coverny2
```

Let's now check and see how each layer plots, including the ebird data
```{r, message=F, warning=F}
plot(ebird_data)
plot(land_coverny)
plot(tree_coverny2)
plot(precip_coverny2)
plot(elevation_coverny2)

```

Time to stack now that each layer:

```{r, message=F, warning=F}
stack_1= stack(land_coverny,elevation_coverny2,tree_coverny2, precip_coverny2)%>%
  mask(mask=as(ny_counties, "Spatial"))
```


# Step 3: 
```{r, message=F, warning=F}


all_data <-bioclim(stack_1, ebird_data)
stack_1= stack(land_coverny,elevation_coverny2,tree_coverny2, precip_coverny2)%>%
  mask(mask=as(ny_counties, "Spatial"))
e <-extent(-79.7625, -71.77917, 40.475, 45.01667)

data_predict <-predict(all_data, stack_1, progess='text', ext=e)


```

# Results

The first plot shows how each environmental layer is now properly cropped within NYS 
```{r, message=F, warning=F}
plot(stack_1)
```

The Second plot combines all environmental variables and occurnace data taken from ebird
```{r, message=F, warning=F}
plot(data_predict)
```


The third and final plot is a more defined version of the second plot. 
```{r, message=F, warning=F}
gplot(data_predict) + geom_tile(aes(fill = value)) +
  facet_wrap(~ variable) +
  scale_fill_viridis(option="turbo") +
  coord_equal()+ ylab("Latitude")+ xlab("Longitude")+ggtitle("How Environmental Variables effect Turkey populations within NYS")

```



# Conclusions


# References

“Global Forest Change 2000–2020data Download.” Global Forest Change, https://storage.googleapis.com/earthenginepartners-hansen/GFC-2020-v1.8/download.html. 

PRISM Climate Group, Oregon State U. (n.d.). Prism.oregonstate.edu. https://prism.oregonstate.edu/recent/monthly.php

USGS (1997) USGS 30 ARC-second Global Elevation Data, GTOPO30. (ed. C.a.I.S.L. Research Data Archive at the National Center for Atmospheric Research).


