---
title: "How landscape habitat influences wild eastern turkey populations within New York State"
author: Marissa Zimmer 
output:
    html_document

---

# Introduction

New York State’s (NYS) wild turkey population constitutes a significant part of conservation concerns and support for the hunting community within NYS. In the early 1950s, the Department of Conservation  (NYSDEC) helped restore and repopulate NYS's wild turkey populations. The NYSDEC also partnered with the National Wild Turkey Federation and SUNY Environmental Science and Forestry college in 2010 to research the impact of weather patterns and hunting zones on turkey populations. The results of this study have not been published at this time. The aim of this project is to research, gather and compare data that involves understanding and explaining how land cover and tree canopy cover can affect wild turkey population levels. Within NYS many individuals have noticed seeing less wild turkeys than in years past and changing habitat factors in localized areas may be one factor affecting population levels. By looking at land cover and tree canopy cover the aim of this project is to see if areas with less suitable land cover and tree canopy cover correlates to areas of lower population levels. If the results of this study come back to support this hypothesis then there is room for continued research into habitat suitability within NYS for wild turkey populations. 



# Materials and methods


Narrative: For this project there were four environmental factors that were used; landcover, treecover, elevation, and precipitation.

*Landcover-USGS 
*Treecover-Global Forest Watch
*Elevation-Dr. Williams 
*Precipitation-PRISM 



See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}
1+2
```

Load any required packages in a code chunk (you may need to install some packages):

```
library(raster)
library(sf)
library(tigris)
library(rasterVis)
library(rgdal)
library(dismo)


options(tigris_use_cache = TRUE)
  # cache the results for quick compiling
```

## Download and clean all required data

land_cover = raster ("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/landcover_crop.asc")

tree_cover = raster("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/Hansen_GFC-2019-v1.7_treecover2000_50N_080W (1).tif")

precip_cover = raster("C:/USers/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/PRISM_ppt_stable_4kmM3_2020_asc.asc")

elevation_cover = raster ("C:/Users/maris/OneDrive/Desktop/Environmental Factor for Turkey Project/elevation.tif")


#Getting all environmental factors on the same resolution, extent: lat/long, crs:4326

ny_counties <-counties(state = "NY", cb = FALSE,
                       resolution = "500k", year = 2020) %>%
                        st_union() %>%
                         st_transform(crs=projection(land_cover))

land_coverny= crop(land_cover, as(ny_counties, "Spatial"))

tree_coverny = crop(tree_cover, 
                    as(st_transform(ny_counties,st_crs(tree_cover)), "Spatial"))#%>%

tree_coverny2 = projectRaster(tree_coverny, to=land_coverny,method="ngb")

elevation_coverny= crop(elevation_cover,
                   as(st_transform(ny_counties,st_crs(elevation_cover)), "Spatial"))#%>%

elevation_coverny2 = projectRaster(elevation_coverny,to=land_coverny, method = "ngb")

precip_coverny = crop(precip_cover, 
                      as(st_transform(ny_counties,st_crs(precip_cover)), "Spatial"))#%>%

precip_coverny2 = projectRaster(precip_cover, to=land_coverny,method="ngb")

#masking, stacking, creating my map.

extent_1 = extent(land_coverny)
precip_cover = crop(precip_cover, extent_1)
tree_coverny2 = crop(tree_coverny2, extent_1)
elevation_coverny2 = crop(elevation_coverny2, extent_1)

stack_1= stack(land_coverny,elevation_coverny2,tree_coverny2, precip_coverny2)
plot(stack_1)

# Results


Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner

